SPP-Perc: Peceptron Learning Based Enhanced Signature Path Prefetcher

### 2018-12-06 back.tex
Mostly related to images, methodology, typos etc

While the basic idea of perceptron based prefetch filtering is applicable to
any look-ahead prefetcher, we develop a practical implementation of the idea
using SPP as our baseline.  Here we describe the basic architecture of SPP.

\textbf{Signature Table:} The Signature Table keeps track of 256 most recently
accessed pages.  It is meant to capture memory access patterns within a page
boundary.  SPP indexes into an entry of the Signature Table using the page
number.  For each entry corresponding to a page, the table stores a `last
block offset' and an `old signature'.  Last block offset is the block offset
of the last memory access of that given page.  The block offset is calculated
with respect to the page boundary.  The signature is a 12-bit compressed
representation of the past few memory accesses for that page.  The signature
is calculated as: $$New Signature = (\,Old Signature << 3 bits\,) \;\;XOR\;\;
(\,Delta\,)$$ Delta is the numerical difference between the block offset of
the current and the previous memory access. In case a matching page entry is
found, the stored signature retrieved and used to index into the Pattern
Table. This process is illustrated by figure \ref{fig:spp_update}.

\textbf{Pattern Table:} The Pattern Table is indexed by the signature
generated from the Signature Table.  Pattern Table holds predicted delta
patterns and their confidence estimates. Each entry indexed by the signature
holds up to 4 unique delta predictions.  This is implemented by making the
Pattern Table 4-way associative.

@@

\textbf{Prefetch Table:} The Prefetch Table is a 1024 entry 1-way associative
table that keeps a record of last few entries prefetched.  This proves useful
for updating the prefetcher states if a tracked prefetch leads to a demand hit
or a cache eviction. \textit{Note:} The original paper refers to this as the
Prefetch Filter but since in our context filter refers to the perceptron
filter, we will be calling this structure as the Prefetch Table to avoid any
confusion.

\textbf{Lookahead Prefetching:} On each trigger, SPP goes down the program
speculation path using its own prefetch suggestion.  Using current prefetch as
the baseline, it re-accesses the Pattern Table to generate further prefetches.
It repeats the cycle of accessing Pattern Table and updating the signature
based on highest confidence prefetch from the last iteration.  The iteration
counter on which SPP manages to predict prefetch entries in the lookahead
manner is characterized as its `depth'.  While doing so, SPP also keeps
compounding the confidence in each depth.  Thus as depth increases, overall
confidence keeps decreasing.

\textbf{Confidence Tracking}: The Pattern Table keeps a count of hits to each
signature through a counter C\textsubscript{sig}.  The number of hits for a
given delta per signature are tracked using a counter C\textsubscript{delta}.
The confidence for a given delta is approximated through C\textsubscript{d} =
C\textsubscript{delta} / C\textsubscript{sig}.  When SPP enters into a
lookahead mode, the path confidence P\textsubscript{d} is given as:
$$P\textsubscript{d} \;=\; \alpha  \;.\;  C\textsubscript{d}  \;.\;
P\textsubscript{d-1}$$ Here $\alpha$ represents the global accuracy,
calculated as the ratio of the number of prefetches which led to a demand hit
to the number of prefetches recommended in total. The range of $\alpha$ is
[0,1].

`D' is the lookahead depth. For $d = 1$, when SPP is in
non-speculative mode, P\textsubscript{0} can be thought of as 1. 
The final P\textsubscript{d} is thresholded against PF\_THRESHOLD 
(T\textsubscript{p}) to reject the low confidence suggestions and 
then against a numerically bigger FILL\_THRESHOLD (T\textsubscript{f}) to 
decide whether the prefetch should be sent to L2 Cache 
(high confidence prefetch) or Last Level Cache (low confidence prefetch).
The two thresholds were empirically set to 25 and 90 respectively, 
on the scale of 0 to 100.

\begin{figure}
  \begin{center}
  \includegraphics[width=9cm]{SPP_Prefetch_Description}
    \label{fig:spp_strcture}
  \caption{SPP architecture}
  \end{center}
\end{figure}

\begin{figure}
  \begin{center}
  \includegraphics[width=9cm]{SPP_Update_Description}
  \label{fig:spp_update}
  \caption{SPP data-path flow}
  \end{center}
\end{figure}

>>>>

While the basic idea of perceptron based prefetch filtering is applicable to
any look-ahead prefetcher, we develop a practical implementation of the idea
using SPP as our baseline. Here we describe the basic architecture of SPP.\\

\noindent \textbf{Signature Table:} The Signature Table keeps track of 256 most recently
accessed pages. It is meant to capture memory access patterns within a page
boundary. SPP indexes into an entry of the Signature Table using the page
number. For each entry corresponding to a page, the table stores a `last
block offset' and an `old signature'. Last block offset is the block offset
of the last memory access of that given page. The block offset is calculated
with respect to the page boundary. The signature is a 12-bit compressed
representation of the past few memory accesses for that page. The signature
is calculated as: $$New Signature = (\,Old Signature << 3 bits\,) \;\;XOR\;\;
(\,Delta\,)$$ Delta is the numerical difference between the block offset of
the current and the previous memory access. In case a matching page entry is
found, the stored signature retrieved and used to index into the Pattern
Table. This process is illustrated by figure \ref{fig:spp_update}.\\

\noindent \textbf{Pattern Table:} The Pattern Table is indexed by the signature
generated from the Signature Table. Pattern Table holds predicted delta
patterns and their confidence estimates. Each entry indexed by the signature
holds up to 4 unique delta predictions. This is implemented by making the
Pattern Table 4-way associative.\\

@@

\noindent \textbf{Prefetch Table:} The Prefetch Table is a 1024 entry 1-way associative
table that keeps a record of last few entries prefetched. This proves useful
for updating the prefetcher states if a tracked prefetch leads to a demand hit
or a cache eviction. \textit{Note:} The original paper refers to this as the
Prefetch Filter but since in our context filter refers to the perceptron
filter, we will be calling this structure as the Prefetch Table to avoid any
confusion.\\

\noindent \textbf{Lookahead Prefetching:} On each trigger, SPP goes down the program
speculation path using its own prefetch suggestion. Using current prefetch as
the baseline, it re-accesses the Pattern Table to generate further prefetches.
It repeats the cycle of accessing Pattern Table and updating the signature
based on highest confidence prefetch from the last iteration. The iteration
counter on which SPP manages to predict prefetch entries in the lookahead
manner is characterized as its `depth'. While doing so, SPP also keeps
compounding the confidence in each depth. Thus as depth increases, overall
confidence keeps decreasing.\\

\noindent \textbf{Confidence Tracking}: The Pattern Table keeps a count of hits to each
signature through a counter C\textsubscript{sig}. The number of hits for a
given delta per signature are tracked using a counter C\textsubscript{delta}.
The confidence for a given delta is approximated through C\textsubscript{d} =
C\textsubscript{delta} / C\textsubscript{sig}. When SPP enters into a
lookahead mode, the path confidence P\textsubscript{d} is given as:
$$P\textsubscript{d} \;=\; \alpha  \;.\;  C\textsubscript{d}  \;.\;
P\textsubscript{d-1}$$ Here $\alpha$ represents the global accuracy,
calculated as the ratio of the number of prefetches which led to a demand hit
to the number of prefetches recommended in total. The range of $\alpha$ is [0,1].
$D$ is the lookahead depth. For $d = 1$, when SPP is in
non-speculative mode, P\textsubscript{0} can be thought of as 1. 
The final P\textsubscript{d} is thresholded against prefetch threshold 
(T\textsubscript{p}) to reject the low confidence suggestions and 
then against a numerically bigger fill threshold (T\textsubscript{f}) to 
decide whether the prefetch should be sent to L2 Cache 
(high confidence prefetch) or Last Level Cache (low confidence prefetch).
The two thresholds were empirically set to 25 and 90 respectively, 
on the scale of 0 to 100.

\begin{figure}
  \begin{center}
  \includegraphics[width=\columnwidth]{SPP_Prefetch_Description}
    \label{fig:spp_strcture}
  \caption{SPP Architecture}
  \end{center}
\end{figure}

\begin{figure}
  \begin{center}
  \includegraphics[width=\columnwidth]{SPP_Update_Description}
  \label{fig:spp_update}
  \caption{SPP Data-path Flow}
  \end{center}
\end{figure}

### 2018-12-06 design_arch.tex
Mostly related to images, methodology, typos etc

\begin{figure}
  \begin{center}
  \includegraphics[width=8cm]{PPF_Hierarchy.pdf}
    \label{fig:PPF_Hierarchy}
  \caption{PPF Architecture in the Memory Hierarchy}
  \end{center}
\end{figure}

It can be beneficial to allow a prefetcher to speculate as deeply as
possible. Often, some useful prefetches are generated long after the
confidence of the prefetcher has fallen below a point, after which the
performance degrades, due to the increase of inaccurate prefetches.  However,
in order to allow deep speculation in the prefetcher, inaccurate prefetches
must be filtered out.  We propose to leverage perceptron-based learning as a
mechanism to differentiate between potentially useful deeply speculated
prefetches and likely not-useful ones. The Perceptron Prefetch Filter (PPF) is
placed between the prefetcher and the prefetch insertion queue, as illustrated
in figure ~\ref{fig:PPF_Hierarchy}, to prevent not-useful prefetches from
polluting the higher levels of the memory hierarchy.

@@

\subsection{The Perceptron Filter}
\label{Arch-Perceptron}
\pg{rewrite this section as talking to the figures and walking through
  them}

Figure~\ref{fig:PPF_Datapath} shows the required structures of PPF, as well as
the steps followed to filter out not-useful prefetches.  The filter is
organized as a set of tables, where each entry in the tables holds a
\textit{weight}. For a configuration of PPF using $N$ number of features, $N$
different tables of weights are needed. Each feature is used to index a
distinct table. The number of entries of each table varies according to their
corresponding feature, hence, different number of bits are needed to index
different tables.  Each weight is a 5-bit saturating counter ranging from -16
to +15.  We found that having 5-bit weights provide a good trade-off between
accuracy and area. A detailed explanation of the storage overhead of PPF can
be found in Figure~\ref{tab:PPF_overhead}.

\begin{figure}
  \begin{center}
  \includegraphics[width=9cm]{Datapath.pdf}
    \label{fig:PPF_Datapath}
  \caption{PPF Data Path and Operation}
  \end{center}
\end{figure}

\noindent{\textbf{Inference}\\
%ETG:is this like an explanation of how prefetching happens? 
%EB: I see this as a short context of the datapath leading to the perc filter.
The base prefetcher is triggered on every demand access to the L2 Cache.  At
this point, it has the opportunity to trigger a prefetch. If it decides to do
so, it will also need to decide how many cache blocks to prefetch.  These
blocks can be either placed in the L2 or L3 cache according to the confidence
of the prefetching mechanism.  Once the base prefetcher is triggered, the
suggested prefetch candidates are fed to the perceptron filter to determine
the usefulness of these prefetches.  The filter makes the decision to issue
the prefetch suggestions from the base prefetcher.  To make the decision, each
feature corresponding to a suggested prefetch is used to index a table and all
the corresponding weights are summed.  The sum denotes the confidence value
for the suggested prefetch, and is thresholded against two different
magnitudes: PERC\_THRESHOLD\_HI and PERC\_ THRESHOLD\_ LO.

Prefetches whose sum exceeds PERC\_THRESHOLD\_HI are placed into the L2 cache.
The higher confidence value hints the prefetch would be useful and should be
prioritized.  A prefetch for which the features result in a confidence value
between PERC\_THRESHOLD\_LO and PERC\_THRESHOLD\_HI is allocated in the larger
LLC, as the filter is moderately confident of the future reuse of the cache
block, but not enough to possible pollute a significantly smaller L2.
Suggested prefetches for which the features lead to a confidence value lower
than PERC\_THRESHOLD\_LO are not prefetched, as the low confidence value
represents that the perceptron has learn that similar set of features are
associated with non-useful prefetches. \newline

\noindent{2. Recording}\\
As shown in step 2 of figure ~\ref{fig:PPF_Datapath} the prefetches that make
it through the inference stage are recorded in the ``Prefetch Table''. The
prefetch table is a 1,024-entry, direct mapped structure that contains all
metadata required to re-index the perceptron entries for training.

@@

In general, PPF can be adapted to a new base prefetcher with only a
few modifications: \\ \\
%
\textbf{Enhancing the Base Prefetcher:} By tuning down any internal thresholds
to increase its inherent aggressiveness. \\
%
\textbf{Inferencing and Storing:} All prefetch recommendations are tested
using the perceptron inferencing algorithm. The perceptron's output,
\textit{true} or \textit{false}, should be saved appropriately, along with all
metadata required for perceptron indexing. \\
%
\textbf{Retrieving and Training:} When feedback for a prefetch is available,
the previously stored metadata can be used to re-index into the perceptron
entries and increment or decrement the weights. \\
%
\textbf{Feature Selection:}  Perceptrons essentially normalize contributions
from different features to get a single sum representing the final confidence.
So the perceptron learning can be as good as the set of features chosen.
Interestingly, this is what makes perceptron learning scalable, as it can
easily learn to incorporate newer information in the form of new features.
Some of the features we developed use information derived directly from
program execution, agnostic to the baseline prefetcher. Beyond that, the
feature set can be expanded to convey any useful information from the
prefetcher to the perceptron filter.

>>>>

\begin{figure}
  \begin{center}
  \includegraphics[width=8cm]{PPF_Hierarchy}
    \label{fig:PPF_Hierarchy}
  \caption{PPF Architecture in the Memory Hierarchy}
  \end{center}
\end{figure}

It can be beneficial to allow a prefetcher to speculate as deeply as
possible. Often, some useful prefetches are generated long after the
confidence of the prefetcher has fallen below a point, after which the
performance degrades, due to the increase of inaccurate prefetches. However,
in order to allow deep speculation in the prefetcher, inaccurate prefetches
must be filtered out. We propose to leverage perceptron-based learning as a
mechanism to differentiate between potentially useful deeply speculated
prefetches and likely not-useful ones. The Perceptron Prefetch Filter (PPF) is
placed between the prefetcher and the prefetch insertion queue, as illustrated
in figure~\ref{fig:PPF_Hierarchy}, to prevent not-useful prefetches from
polluting the higher levels of the memory hierarchy.

@@

\begin{figure}[ht]
  \begin{center}
  \includegraphics[width=\columnwidth]{Datapath_Separate}
    \label{fig:PPF_Datapath}
  \caption{PPF Data Path and Operation}
  \end{center}
\end{figure}

\subsection{The Perceptron Filter}
\label{Arch-Perceptron}
%\pg{rewrite this section as talking to the figures and walking through them}

Figure~\ref{fig:PPF_Datapath} shows the required structures of PPF, as well as
the steps followed to filter out not-useful prefetches. The filter is
organized as a set of tables, where each entry in the tables holds a
\textit{weight}. For a configuration of PPF using $N$ number of features, $N$
different tables of weights are needed. Each feature is used to index a
distinct table. The number of entries of each table varies according to their
corresponding feature, hence, different number of bits are needed to index
different tables. Each weight is a 5-bit saturating counter ranging from -16
to +15. We found that having 5-bit weights provide a good trade-off between
accuracy and area. A detailed explanation of the storage overhead of PPF can
be found in Figure~\ref{tab:PPF_overhead}.\\

\noindent{\textbf{Inferencing}\\
%ETG:is this like an explanation of how prefetching happens? 
%EB: I see this as a short context of the datapath leading to the perc filter.
The base prefetcher is triggered on every demand access to the L2 Cache. At
this point, it has the opportunity to trigger a prefetch. If it decides to do
so, it will also need to decide how many cache blocks to prefetch. These
blocks can be either placed in the L2 or L3 cache according to the confidence
of the prefetching mechanism. Once the base prefetcher is triggered, the
suggested prefetch candidates are fed to the perceptron filter to determine
the usefulness of these prefetches. The filter makes the decision to issue
the prefetch suggestions from the base prefetcher. To make the decision, each
feature corresponding to a suggested prefetch is used to index a table and all
the corresponding weights are summed. The sum denotes the confidence value
for the suggested prefetch, and is thresholded against two different
magnitudes: PERC\_THRESHOLD\_HI and PERC\_ THRESHOLD\_ LO.

Prefetches whose sum exceeds PERC\_THRESHOLD\_HI are placed into the L2 cache.
The higher confidence value hints the prefetch would be useful and should be
prioritized. A prefetch for which the features result in a confidence value
between PERC\_THRESHOLD\_LO and PERC\_THRESHOLD\_ HI is allocated in the larger
LLC, as the filter is moderately confident of the future reuse of the cache
block, but not enough to possible pollute a significantly smaller L2.
Suggested prefetches for which the features lead to a confidence value lower
than PERC\_THRESHOLD\_LO are not prefetched, as the low confidence value
represents that the perceptron has learn that similar set of features are
associated with non-useful prefetches. \newline

\noindent \textbf{Recording}\\
As shown in step 2 of figure~\ref{fig:PPF_Datapath} the prefetches that make
it through the inference stage are recorded in the ``Prefetch Table''. The
prefetch table is a 1,024-entry, direct mapped structure that contains all
metadata required to re-index the perceptron entries for training.

@@

In general, PPF can be adapted to a new base prefetcher with only a
few modifications:\\
%
\textbf{Enhancing the Base Prefetcher:} By tuning down any internal thresholds
to increase its inherent aggressiveness. \\
%
\textbf{Inferencing and Storing:} All prefetch recommendations are tested
using the perceptron inferencing algorithm. The perceptron's output,
\textit{true} or \textit{false}, should be saved appropriately, along with all
metadata required for perceptron indexing. \\
%
\textbf{Retrieving and Training:} When feedback for a prefetch is available,
the previously stored metadata can be used to re-index into the perceptron
entries and increment or decrement the weights. \\
%
\textbf{Feature Selection:} Perceptrons essentially normalize contributions
from different features to get a single sum representing the final confidence.
So the perceptron learning can be as good as the set of features chosen.
Interestingly, this is what makes perceptron learning scalable, as it can
easily learn to incorporate newer information in the form of new features.
Some of the features we developed use information derived directly from
program execution, agnostic to the baseline prefetcher. Beyond that, the
feature set can be expanded to convey any useful information from the
prefetcher to the perceptron filter.

### 2018-12-06 method.tex
Mostly related to images, methodology, typos etc

\textit{Single-core performance:} For single-core simulations, we use 200
million instructions to warm-up the microarchitectural structures and the next
one billion instructions to do detailed simulations and collect run-time
statistics. We report the IPC speedup over the baseline of no prefetching.
The final number reported is the geometric mean of the speedup achieved on
individual traces.

\textit{Multi-core performance:} For multi-application workloads, we generate
100 random mixes and another 100 mixes from the memory intensive subset of
SPEC CPU 2017.  For 4-core workload, 200 Million instructions are used for warm-up
and additional 1 Billion instruction simulated for collecting statistics.
Each CPU keeps executing its workload till the last CPU completes one billion
instructions after warm-up.  For collecting IPC and other data, only the first
billion instructions are considered as the region of interest.

@@

We repeat the same process for 8-core workloads, correspondingly with 16MB
LLC.  The only difference is that 20 million warm-up instructions and 100
million full instructions are executed.  This is done so as to keep the
simulation run-time within reasonable limits as a single 8-core mix takes up
to 3 days to simulate one billion instructions.

\textit{Validation:} We cross-validated our PPF model using SPEC
CPU 2006~\cite{SPEC2006} and CloudSuite~\cite{CloudSuite} benchmarks.  For
single-core SPEC CPU 2006, we developed 94 simpoints spread across all the 29
applications. For multi-core, we followed the same methodology as SPEC CPU 2017.
For CloudSuite, we used the traces made available for the 2nd Cache
Replacement Competition (CRC-2)~\cite{CRC_2}.  The traces include 4 4-core
applications with 6 distinct phases per application.

@@

\subsection{Prefetchers Simulated}
\label{Method-Prefetchers}
We compared PPF against three of the latest state of the art hardware-only
prefetchers: Best Offset Prefetcher (BOP), DRAM Aware - Access Map Pattern
Matching (DA-AMPM)~\cite{DA_AMPM} and Signature Path Prefetcher (SPP).  BOP was
the winner of 2nd Data Prefetching Championship.  DA-AMPM is the enhanced
version of AMPM, modified to account for DRAM row buffer locality.  SPP has
been shown to outperform BOP on SPEC CPU 2006 traces.  For each of these, we
compare their speedups taking no prefetching as the baseline.

@@

We plot a histogram for each feature depicting weights distribution from -16
to +15 and generate the Pearson's correlation factor for that feature.
Pearson's factor is a numerical measure ranging from -1 to 1 of the degree of
linear correlation between two variables.  The magnitude of Pearson's factor
gives the extent of correlation and the sign indicates whether it is a
positive correlation or a negative correlation.  Values close to 0 suggest a
low correlation or at times, noisy data.  A value of +1/-1 suggests a
perfectly linear positive / negative correlation respectively.
Figure~\ref{Fig:Hist} depicts the histogram distribution of trained weights,
for a feature with high and another with moderate Pearson's coefficient.

Figure~\ref{Fig:Pvals} shows all the features used, arranged in the increasing
order of their Pearson's factor.  As can be seen 5 out of the 9
features provide a moderate to high correlation, with the magnitude of
P-value > 0.6.  The single most important feature, 
\textit{Confidence XOR Page address} helps provide
a correlation to prefetch outcome with a factor of 0.90. \\ \\
%
\textbf{Per Trace Correlation} \newline Another important way to look
at the perceptron features is to see how much their contribution
varies across the traces.  Here we give special attention to features
with low P-values in.  Figure~\ref{Pvals_Traces} shows the
variation of P-values for three features : \textit{PC XOR Delta},
\textit{Signature XOR Delta} and 
\textit{PC XOR Depth}; across all the SPEC CPU 2017 traces.  
For simplicity, the traces are arranged in an
increasing order of contribution made by the feature and the trace
names have been omitted.  It can be seen that even features with a low
overall correlation provide useful correlation (magnitude > 0.5) for
a significant number of traces. \\ \\
%
\textbf{Trimming Features Using Correlation} \newline Beside
providing interesting insights into prefetching behavior, P-value can
also be used for feature selection and prefetcher tuning.

As we examine correlation of each feature with the final outcome, we also
study correlation between the features.  We used the above methodology to
eliminate features providing little information that has already been captured
in other features.

We initially came up a mix of 23 features- primary and composite.  By studying
cross correlation of each of these features against others in a 23$\times$23
matrix, we identified pairs of features with correlation factor > 0.9 in
magnitude and eliminated redundant features, reducing the feature count to 9.
Thus, in the final implementation of PPF, no two features have high
correlation between them.  This way we can be sure that each feature makes a
contribution that cannot be captured using other features.

Secondly, studying the relative importance of each feature enabled us to vary
the number of entries dedicated for each feature.  Features with higher
correlation, \textit{cache line} and \textit{page address} were given most
importance and allowed full 12-bits of indexing.  Features like \textit{IP XOR
delta} and \textit{IP XOR depth} with a low overall P-value were allocated
fewer entries in the feature table.

@@

\begin{table}[ht]
    \centering
    \begin{tabular}{|c|c|c|}
    \hline
        \textbf{Field} &
        \textbf{Bits} &
        \textbf{Comment} \\
    \hline
         Valid 		& 1  & Indicates a valid entry in the table\\
         Tag 		& 6  & Identifier for the entry in the table\\
         \multirow{2}{1cm}{Useful} 	& 1  & To show if the given entry led to a\\
                    	&    & useful demand fetch\\
         Perc Decision 	& 1  & Prefetched vs Not-prefetched \\
    \hline
        PC 		& 12 & \multirow{4}{4.8cm}{Meta-data required for perceptron}\\
        Address 	& 24 & \multirow{4}{1.1cm}{training}\\
        Delta 		& 7  & \\
        Curr Signature 	& 12 & \\
        Confidence 	& 7  & \\
    \hline
        \multicolumn{3}{|c|}{\textbf{Total 84 bits}}\\
    \hline
    \end{tabular}
    \caption{Meta-data Stored in Prefetch Table}
    \label{tab:PTable_metadata}
\end{table}

>>>>

\noindent \textit{Single-core performance:} For single-core simulations, we use 200
million instructions to warm-up the microarchitectural structures and the next
one billion instructions to do detailed simulations and collect run-time
statistics. We report the IPC speedup over the baseline of no prefetching.
The final numbers reported are the geometric mean of the weighted mean speedup 
achieved per application using the SimPoint methodology.\\

\noindent \textit{Multi-core performance:} For multi-application workloads, we generate
100 random mixes and another 100 mixes from the memory intensive subset of
SPEC CPU 2017. For 4-core workload, 200 Million instructions are used for warm-up
and additional 1 Billion instruction simulated for collecting statistics.
Each CPU keeps executing its workload till the last CPU completes one billion
instructions after warm-up. For collecting IPC and other data, only the first
billion instructions are considered as the region of interest.\\

@@

We repeat the same process for 8-core workloads, correspondingly with 16MB
LLC. The only difference is that 20 million warm-up instructions and 100
million full instructions are executed. This is done so as to keep the
simulation run-time within reasonable limits as a single 8-core mix takes up
to 3 days to simulate one billion instructions.\\

\noindent \textit{Validation:} We cross-validated our PPF model using SPEC
CPU 2006~\cite{SPEC2006} and CloudSuite~\cite{CloudSuite} benchmarks. For
single-core SPEC CPU 2006, we developed 94 simpoints spread across all the 29
applications. For multi-core, we followed the same methodology as SPEC CPU 2017.
For CloudSuite, we used the traces made available for the 2nd Cache
Replacement Competition (CRC-2)~\cite{CRC_2}. The traces include 4 4-core
applications with 6 distinct phases per application.

@@

\subsection{Prefetchers Simulated}
\label{Method-Prefetchers}
We compared PPF against three of the latest state of the art hardware-only
prefetchers: Best Offset Prefetcher (BOP), DRAM Aware - Access Map Pattern
Matching (DA-AMPM) \cite{DA_AMPM} and Signature Path Prefetcher (SPP). BOP was
the winner of 2nd Data Prefetching Championship. DA-AMPM is the enhanced
version of AMPM, modified to account for DRAM row buffer locality. SPP has
been shown to outperform BOP on SPEC CPU 2006 traces. For each of these, we
compare their speedups taking no prefetching as the baseline.

@@

We plot a histogram for each feature depicting weights distribution from -16
to +15 and generate the Pearson's correlation factor for that feature.
Pearson's factor is a numerical measure ranging from -1 to 1 of the degree of
linear correlation between two variables. The magnitude of Pearson's factor
gives the extent of correlation and the sign indicates whether it is a
positive correlation or a negative correlation. Values close to 0 suggest a
low correlation or at times, noisy data. A value of +1/-1 suggests a
perfectly linear positive / negative correlation respectively.

As a part of our perceptron feature selection methodology, we explored a wide 
variety of features to begin with. Features with a low 
Pearson's coefficient were rejected as they didn't provide much useful 
correlation. Figure~\ref{Fig:Hist} depicts the histogram distribution of 
trained weights for two features. The first feature, 
\textit{Confidence XOR Page address} has the highest observed P-value, 
hence was retained. On the other hand, the second feature, \textit{Last Signature}
did not provide any meaningful correlation and hence was rejected. 
This is visible from the bulk of its trained weights settling to near zero values.

Figure~\ref{Fig:Pvals} shows all the features which are finally used, 
arranged in the increasing order of their Pearson's factor. 
As can be seen 5 out of the 9 features provide a moderate to high 
correlation, with the magnitude of P-value > 0.6. The single most 
important feature, \textit{Confidence XOR Page address} helps provide
a correlation to prefetch outcome with a factor of 0.90. \\ \\
%
\textbf{Per Trace Correlation} \newline Another important way to look
at the perceptron features is to see how much their contribution
varies across the traces. Here we give special attention to features
with low P-values in. Figure~\ref{Fig:Pvals_Trace} shows the
variation of P-values for three features : \textit{PC XOR Delta},
\textit{Signature XOR Delta} and 
\textit{PC XOR Depth}; across all the SPEC CPU 2017 traces. 
For simplicity, the traces are arranged in an
increasing order of contribution made by the. 
It can be seen that even features with a low
overall correlation provide useful correlation (magnitude > 0.5) for
a significant number of traces. \\ \\
%
\textbf{Trimming Features Using Cross-correlation} \newline Beside
providing interesting insights into prefetching behavior, P-value can
also be used for feature selection and prefetcher tuning.

As we examined correlation of each feature with the final outcome, we also
studied correlation between the features. We used the above methodology to
eliminate features providing little information that has already been captured
in other features.

We initially came up a mix of 23 features. By studying
cross correlation of each of these features against others in a 23$\times$23
matrix, we identified pairs of features with correlation factor > 0.9 in
magnitude and eliminated redundant features, reducing the feature count to 9.
Thus, in the final implementation of PPF, no two features have a high
correlation between them. This way we can be sure that each feature makes a
contribution that cannot be captured using other features.

Secondly, studying the relative importance of each feature enabled us to vary
the number of entries dedicated for each feature. Features with higher
correlation, \textit{cache line} and \textit{page address} were given most
importance and allowed full 12-bits of indexing. Features like \textit{PC XOR
delta} and \textit{PC XOR depth} with a low overall P-value were allocated
fewer entries in the feature table.

@@

\begin{table}[ht]
    \centering
    \begin{tabular}{|c|c|c|}
    \hline
        \textbf{Field} &
        \textbf{Bits} &
        \textbf{Comment} \\
    \hline
         Valid 		& 1  & Indicates a valid entry in the table\\
         Tag 		& 6  & Identifier for the entry in the table\\
         \multirow{2}{1cm}{Useful} 	& 1  & To show if the given entry led to a\\
                    	&    & useful demand fetch\\
         Perc Decision 	& 1  & Prefetched vs Not-prefetched \\
    \hline
        PC 		& 12 & \\
        Address 	& 24 & \\
        Curr Signature 	& 10 & \multirow{2}{4.8cm}{Meta-data required for perceptron}\\
	PC$_i$ Hash	& 12 & \multirow{2}{1.1cm}{training}\\
        Delta 		& 7  & \\
        Confidence 	& 7  & \\
	Depth		& 4  & \\
    \hline
        \multicolumn{3}{|c|}{\textbf{Total 85 bits}}\\
    \hline
    \end{tabular}
    \caption{Meta-data Stored in Prefetch Table}
    \label{tab:PTable_metadata}
\end{table}
